---
title: "viz"
author: "Nina Bingham"
date: "April 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Now we are going to work on visualization

``` {r load, results='hide', message=FALSE, warning=FALSE}

## install ggplot2 and tmap ...only need to do once, did on april 16th, 2016!
# install.packages("ggplot2")
# install.packages("tmap")

library("dplyr")
library("gapminder") #here we loaded the dplyr library first and because of that when we loaded gapminder is presented a pretty nice looking dataframe.
library("ggplot2")
library("tmap") 
```

## Scatter Plots n stuff

``` {r}

# setup dataframe
g = gapminder %>%
  filter(year==2007) %>%   # most recent year 
  mutate(pop_m = pop/1e6)  # population, millions

# plot scatterplot of most recent year 
s = ggplot(g, aes(x=gdpPercap, y=lifeExp)) +
  geom_point()
s # when it actuall puts it out

# aes is saying.. what do you want to assign for your x and y axis here?
# geom_point is adding the points in, look in cheatsheets for type of data

# you can add on different data styles to stuff...
s= s + 
  aes(size=pop_m) + aes(color=continent)

# add title, update axes labels
s = s + 
  ggtitle('Health & Wealth of Nations for 2007') +
  xlab('GDP per capita ($/year)') +
  ylab('Life expectancy (years)')

# label legend
s = s + 
  scale_colour_discrete(name='Continent') +
  scale_size_continuous(name='Population (M)')

s

ggsave('hans.png')
```

## now let's make box plots

```{r}

# boxplot by continent
b = ggplot(g, aes(x=continent, y=lifeExp)) +
  geom_boxplot() + aes(fill=continent) + 
  theme(legend.position='none') +
  ggtitle('Life Expectancy by Continent for 2007') +
  xlab('Continent') +
  ylab('Life expectancy (years)')
b

#going to plotly to make it INTERACTIVE

#install.packages('plotly') this was done april 16th 2016
library(plotly) 

# scatterplot (Note: key=country shows up on rollover)
s = ggplot(g, aes(x=gdpPercap, y=lifeExp, key=country)) +
  geom_point()
ggplotly(s)